---
title: "COVID-19 webpage"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
    theme: cosmo
runtime: shiny
resource_files:
- TM_WORLD_BORDERS_SIMPL-0.3.dbf
- TM_WORLD_BORDERS_SIMPL-0.3.prj
- TM_WORLD_BORDERS_SIMPL-0.3.shp
- TM_WORLD_BORDERS_SIMPL-0.3.shx
---

```{r setup, include=FALSE}
library(flexdashboard)
library(rvest)
library(xml2)
library(stringr)
library(leaflet)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(ggiraph)
library(plotly)
library(usmap)
library(shiny)
library(tidyverse)
library(magrittr)
library(reshape2)
library(countrycode)
library(rworldmap)
library(rgdal)
library(rgeos)
library(devtools)
library(tweenr)
library(curl)
library(viridis)
library(maps)
library(plyr)
```

Country with travel restriction {.sidebar}
================================================================

```{r}
a = Sys.time()
a = .POSIXct(a, "EST5EDT")
renderText(paste('Last Updated on:',a, 'EDT'))
```

### Country with travel restriction  

These countries are listed with Level 3 Travel Health Notice by the CDC:

```{r}
# Download world information
xhw = xml2::read_html("https://www.worldometers.info/coronavirus/")
tcounts = xhw %>%
  html_nodes('#maincounter-wrap span') %>%
  html_text() %>%
  str_remove_all(' ') %>%
  str_remove_all(',') %>%
  as.numeric()
  
# Country information
countrytable = xhw %>%
  html_nodes('td') %>%
  html_text() %>%
  str_remove_all(' ') %>%
  str_remove_all(',') %>%
  str_remove_all('\n')
rows = length(countrytable)/9
countrytable = countrytable %>%
  matrix(9,rows) %>%
  t()
countrytable = countrytable[1:rows-1,]
countrytable = as.data.frame(countrytable, stringsAsFactors=FALSE)
countrytable = countrytable[1:which(grepl('Total:',countrytable$V1))-1,]

# Lat and Lng info for each country
countries = xml2::read_html("https://developers.google.com/public-data/docs/canonical/countries_csv")
countriesloc = countries %>%
  html_nodes('td') %>%
  html_text() %>%
  str_remove_all(' ') %>%
  matrix(4,245) %>%
  t() %>%
  as.data.frame(stringsAsFactors=FALSE)
colnames(countriesloc) = c('abb','lat','lng','cname')
countriesloc[228,'cname'] = 'USA'
countrytable = countrytable[countrytable[['V1']] %in% countriesloc[['cname']], ]
countrydata = merge(countriesloc, countrytable, by.x='cname',by.y='V1')

# List all level 3 countries
level3 = xml2::read_html("https://www.cdc.gov/coronavirus/2019-ncov/travelers/after-travel-precautions.html")
l3 = level3 %>%
  html_nodes('.col-md-6 li') %>%
  html_text() %>%
  as.data.frame()
renderTable(l3, colnames=FALSE)

# Codes for United States
#load usa coronavirus data
usadat <- read_html("https://www.worldometers.info/coronavirus/country/us/")
usadf <- html_table(usadat,fill=TRUE)[[1]]

# buid a new usa dataframe 
state <- usadf$USAState
number <- length(state)
state <- state[-number] 
Totalcase <- usadf$TotalCases
Totalcase <- Totalcase[-number]
Totalcase <- as.numeric(gsub(",","",Totalcase))
usadfnew <- data.frame(state,Totalcase)
usadfnew$hover <- with(usadfnew, paste(state, '<br>', "case number:", Totalcase))

#combine state code 
dfstatecode <- read.csv("https://raw.githubusercontent.com/jasonong/List-of-US-States/master/states.csv")
usadfnew$code <- dfstatecode$Abbreviation[match(usadfnew$state, dfstatecode$State)]

# light grey boundaries
l <- list(color = toRGB("white"), width = 2)

# specify map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
```

World
================================================================

Row {data-height=180}
----------------------------------------------------------------

### Total Cases {.value-box}

```{r}
renderValueBox({
  valueBox(
    value = tcounts[1],
    icon = "fa-globe-asia"
  )
})
```

### Total Deaths {.value-box}

```{r}
renderValueBox({
  valueBox(
    value = tcounts[2],
    icon = "fa-procedures"
  )
})
```

### Total Recovered {.value-box}

```{r}
renderValueBox({
  valueBox(
    value = tcounts[3],
    icon = "fa-notes-medical"
  )
})
```

Row {.tabset .tabset-fade}
----------------------------------------------------------------

### Total COVID-19 diagnosed by Country (click circles for numbers)

```{r}
renderLeaflet({
  map = leaflet(countrytable) %>%
    addTiles() %>%
    setView(2.530401, 36.020884, zoom = 2) %>%
    clearShapes() %>%
    addCircles(lat = as.numeric(countrydata[['lat']]), lng = as.numeric(countrydata[['lng']]), radius = as.numeric(countrydata[['V2']]) * 20, color = 'Red', fill = TRUE, popup=paste(countrydata[['cname']],' total cases: ',countrydata[['V2']],sep=''))
  map
})
```

### Active cases by country

```{r}
# Get active cases
COVID_data <- xml2::read_html("https://www.worldometers.info/coronavirus/", stringsAsFactors=FALSE)
dat <- COVID_data %>%
  html_node("table") %>%
  html_table() 
names(dat)[1] <- c("Country")
dat <- apply(dat, 2, function(x) gsub("^$|^ $", NA, x))
dat[is.na(dat)] <- 0
dat_trans <- as.data.frame(t(dat))
country_name = dat[,1]
names(dat_trans) <- country_name
active <- dat_trans[c("Country","ActiveCases"),]
active <- as.data.frame(t(active))
names(active) <- c("NAME","ActiveCases")
active["ISO3"] <- countrycode(active$NAME, 'country.name', 'iso3c')
active$ActiveCases <- as.numeric(gsub(",","",active$ActiveCases))
active <- filter(active, NAME != c("Total:", "Diamond Princess)"))
world_spdf <- readOGR( 
  dsn= (".") , 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)
world_spdf@data <- join(world_spdf@data, active, 
                         by = "ISO3")
world_spdf@data[,"ActiveCases"][is.na(world_spdf@data[,"ActiveCases"])] <- 0
mytext <- paste(
    "Area: ", world_spdf@data$NAME,"<br/>", 
    "Active Cases: ", round(world_spdf@data$ActiveCases, 2), 
    sep="") %>%
  lapply(htmltools::HTML)
mybins <- c(0,100,200,500,1000,Inf)
mypalette <- colorBin( palette="YlOrBr", domain=world_spdf@data$ActiveCases, na.color="transparent", bins=mybins)

# Plot active cases
map1<-leaflet(world_spdf) %>%
  addTiles() %>%
  setView(lat=10, lng=0 , zoom=2) %>%
  addPolygons(
    fillColor = ~mypalette(ActiveCases), 
    stroke=TRUE, 
    fillOpacity = 0.9, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto")) %>%
  addLegend(pal = mypalette, 
            values = ~ActiveCases, 
            position = "bottomright", 
            title = "COVID-19 Active Cases") 
map1
```

Trends
===============================================================

Row {.tabset .tabset-fade}
-------------------------------------

### Total Confirmed Cases

```{r}
# Data Trends
daily_confirm = read.csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv"))
col_confirm = as.numeric(ncol(daily_confirm))
daily_confirm <- daily_confirm[,5:col_confirm]
names_col <- sub("X", "", colnames(daily_confirm)) 
Dates_confirm <- substr(names_col, 1, 4)
colnames(daily_confirm) <- Dates_confirm
sum_confirm <- as.data.frame(colSums(daily_confirm, dims=1))
sum_confirm  <- sum_confirm  %>% cbind(Dates_confirm)
colnames(sum_confirm) <- c("Number", "Date")
sum_confirm$Date <- factor(sum_confirm$Date, levels = sum_confirm$Date)
daily_death = read.csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv"))
col_death = as.numeric(ncol(daily_death))
daily_death <- daily_death[,5:col_death]
names_col_death <- sub("X", "", colnames(daily_death)) 
Dates_death <- substr(names_col_death, 1, 4)
colnames(daily_death) <- Dates_death
sum_death <- as.data.frame(colSums(daily_death, dims=1))
sum_death  <- sum_death  %>% cbind(Dates_death)
colnames(sum_death) <- c("Number", "Date")
sum_death$Date <- factor(sum_death$Date, levels = sum_death$Date)
max_confirm <- sum_confirm[which.max(sum_confirm$Number),] 
max_confirm <- max_confirm$Number
min_confirm <- sum_confirm[which.min(sum_confirm$Number),] 
min_confirm <- min_confirm$Number
diff_confirm = ceiling((max_confirm - min_confirm) / 5)
minors_confirm <-seq(min_confirm,max_confirm,by=diff_confirm)
max_death <- sum_death[which.max(sum_death$Number),] 
max_death <- max_death$Number
min_death <- sum_death[which.min(sum_death$Number),] 
min_death <- min_death$Number
diff_death = ceiling((max_death - min_death) / 5)
minors_death <-seq(min_death,max_death,by=diff_death)
renderPlotly({
  p <- ggplot(sum_confirm,
             aes(x = Date, y = Number, group=1)) +
             geom_point(shape=23, fill="black", color="black", size=1.5) +
             geom_line(linetype="solid", color="red", size=0.5)+
             ggtitle("Worldwide total confirmed cases by date") +
             geom_hline(mapping=NULL, yintercept=minors_confirm,
                        colour='grey80') +
             expand_limits(x = 0, y = 0) +
             theme(axis.text.x = element_text(angle = 90)) +
             theme(panel.grid.major = element_blank(), panel.grid.minor =             element_blank(),
panel.background = element_blank(),
axis.line = element_line(colour = "grey"), axis.ticks = element_blank(),
plot.title = element_text(color="black", size=14, face ="bold"),
axis.title = element_text(color="black", size=12, face ="bold"))
print(p)
})
```
   
### Total Deaths

```{r}
renderPlotly({
  p <- ggplot(sum_death,
             aes(x = Date, y = Number, group=1)) +
             geom_point(shape=23, fill="black", color="black", size=1.5) +
             geom_line(linetype="solid", color="red", size=0.5)+
             ggtitle("Worldwide total deaths by date") +
             geom_hline(mapping=NULL, yintercept=minors_death,
                        colour='grey80') +
             expand_limits(x = 0, y = 0) +
             theme(axis.text.x = element_text(angle = 90)) +
             theme(panel.grid.major = element_blank(), panel.grid.minor =             element_blank(),
panel.background = element_blank(),
axis.line = element_line(colour = "grey"), axis.ticks = element_blank(),
plot.title = element_text(color="black", size=14, face ="bold"),
axis.title = element_text(color="black", size=12, face ="bold"))
print(p)
})
```

Country
===============================================================
```{r}
renderTable({
  countrydata = countrydata[,-c(2,3,4)]
  countrydata = countrydata[,-c(8,9)]
  countrydata[countrydata==''] = 0
  countrydata[['V2']] = as.numeric(countrydata[['V2']])
  countrydata = countrydata[rev(order(countrydata[['V2']])),]
  countrydata[['V2']] = as.character(countrydata[['V2']])
  colnames(countrydata) = c('Country','Total Cases','New Cases','Total Deaths','New Deaths','Total Recovered','Active cases')
  countrydata
})
```

United States
===============================================================

```{r}
renderPlotly({
  fig <- plot_geo(usadfnew, locationmode = 'USA-states')
  fig <- fig %>% add_trace(
    z = ~Totalcase, color = ~Totalcase, colors = 'Reds',
    text = ~hover, locations = ~code, marker = list(line = l)
  )
  fig <- fig %>% colorbar(title = 'case number')
  fig <- fig %>% layout(
    title = 'Coronavirus case in USA',
    geo = g
  )
  fig
})
```

Resources and Disclaimer
===============================================================

### Resources
CDC COVID-19 website:  
https://www.cdc.gov/coronavirus/2019-nCoV/index.html  

Johns Hopkins CSSE COVID-19 Dashboard:  
https://www.arcgis.com/apps/opsdashboard/index.html#/bda7594740fd40299423467b48e9ecf6  

### Disclaimer
This webpage is part of a course assignment, it is not official health information and should not be used as reference in any circumstances.  

### References
Case numbers:  
https://www.worldometers.info/coronavirus/  

Longitude and Latitude for each country:  
https://developers.google.com/public-data/docs/canonical/countries_csv  